---
title: "Analysis of 2024 U.S. Presidential Election Polling Trends"
subtitle: "A Comparison of Trump and Harris"
author: 
  - Jiaxuan Song
  - Zien Gao
  - Shuheng (Jack) Zhou
thanks: "Code and data are available at: [https://github.com/Jiaxuan-Song/U.S-Election-Prediction-2024.git](https://github.com/Jiaxuan-Song/U.S-Election-Prediction-2024.git)."
date: today
date-format: long
abstract: "This report analyzes the polling trends for Donald Trump and Kamala Harris in the 2024 U.S. Presidential election using data from FiveThirtyEight’s 2024 National Presidential Polls. The study focuses on understanding the changes in support for both candidates over time, across different regions, and how polling methodologies affect their respective support rates. The results show that Trump’s support base remains relatively stable throughout the election period, particularly in Republican-leaning states, while Harris exhibits greater fluctuations, especially during mid-2023, with notable regional differences. Additionally, the impact of different polling methods is explored, showing how online and phone surveys may influence the results. These findings highlight the importance of swing states and the variability across pollsters in shaping election forecasts. The report concludes with a discussion on the factors that might influence the outcome of the election and the need for continuous monitoring of polling trends as election day approaches "
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```


# Introduction

The 2024 U.S. Presidential election is one of the most closely watched political events, with intense competition between candidates. Donald Trump, as a former president, is seeking to reclaim the presidency, while Kamala Harris, the current Vice President, is a prominent figure in the Democratic Party. Polls play a crucial role in predicting election outcomes, offering insight into public support for each candidate. This report analyzes polling data from FiveThirtyEight’s 2024 National Presidential Polls, focusing on the trends in support for Trump and Harris. We will explore how time, regions, and pollsters influence the support rates for both candidates.




# Data {#sec-data}

## Overview

The dataset used for predicting the 2024 U.S. Presidential election is sourced from various public opinion polls conducted by reputable polling organizations throughout the United States. These polls provide detailed information such as pollster names, polling methodology, sample sizes, candidate names, political party affiliations, and percentage support for each candidate, specifically for key figures like Kamala Harris and Donald Trump. Additionally, the dataset includes important poll characteristics such as whether the poll was conducted at a national level, the pollster’s rating, and whether the poll involved hypothetical matchups between candidates.

We utilize the statistical programming language R [@citeR] for data cleaning, preprocessing, and analysis. The dataset is processed to remove inconsistencies and missing data, ensuring it is suitable for forecasting election trends. Following   @tellingstories, we consider he codes used and some graphics .


Overview text

## Measurement
	
The dataset reflects real-world polling data gathered through various methods such as phone surveys, online interviews, and ranked-choice voting reallocation. Each poll entry represents the percentage of respondents supporting a specific candidate at a particular time. Polling organizations, such as TIPP Insights, Quantus Insights, and ActiVote, are included, and the dataset captures details about when and where the poll was conducted, the poll’s methodology, and whether it was a general election or hypothetical scenario.
Data is collected and categorized based on responses related to key candidates (e.g., Kamala Harris and Donald Trump), with percentage values representing the share of voters supporting each candidate at the time of polling. After cleaning and verifying the data for consistency and accuracy, the final dataset will be used to model voting trends leading up to the election.


## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.



Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.








# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


